---
title: "DATA 606 - Project Proposal"
author: "Devin Teran"
output: rmdformats::html_clean
---

### Data Preparation

```{r libraries}
library(tidyverse)
library(dplyr)
library(psych)
```

#### Import Data
```{r setup, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
load(url('http://biostat.mc.vanderbilt.edu/wiki/pub/Main/DataSets/counties.sav'))

state <- read_csv('https://raw.githubusercontent.com/devinteran/DATA606-Project/master/states.csv')
```
#### Clean Data

1. We're not going to need all of the columns for our analysis, so I will remove them now.  
2. I'll remove rows that have NAs in the column democrat or republican.
3. Combine the state region data with the counties data
```{r clean-data}

county_data <- left_join(counties,state,by = c("state" = "State Code"))

county_data <- county_data %>% select(county,state,Region,Division,pop.density,college,income,democrat,republican)

county_data <- county_data %>% filter(is.na(democrat) == FALSE)
county_data <- county_data %>% filter(is.na(republican) == FALSE)

```

### Research question 

**You should phrase your research question in a way that matches up with the scope of inference your dataset allows for.**  
Are income,population density, or % college educated residents predictive of % of county residents voted for the Democratic or Republican Presidential nominee in the 1992 election? Do regions follow specific patterns? 

### Cases 

**What are the cases, and how many are there?**  
Each case represents a single county.  There are 3,114 counties observed in the dataset.
```{r case-count}
nrow(county_data)
```

### Data collection 

**Describe the method of data collection.**  
This data was found on the Department of Biostatistics at Vanderbilt University website and Kaggle.

### Type of study 

**What type of study is this (observational/experiment)?**

This is an observational study.  

### Data Source 

**If you collected the data, state self-collected. If not, provide a citation/link.**

The data can be found at the [Department of Biostatistics at Vanderbilt University website](http://biostat.mc.vanderbilt.edu/wiki/Main/DataSets) and scrolling down to 'U.S. counties and 1992 presidential election dataset'.  The region data was found on [Kaggle](https://www.kaggle.com/omer2040/usa-states-to-region).

### Dependent Variable

**What is the response variable? Is it quantitative or qualitative?**

The response variable is the % of residents that voted for democratic or republican nominee in the 1992 election.  This variable is quantitative.


### Independent Variable

The independent variables are:  

* Pop.density - quantitative
* Income - quantitative  
* % College Educated - quantative 
* Region - qualitative

### Relevant summary statistics 

**Provide summary statistics for each the variables. Also include appropriate visualizations related to your research question (e.g. scatter plot, boxplots, etc). This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**

```{r describe-variables}
describe(county_data$pop.density)
describe(county_data$college)
describe(county_data$income)
describe(county_data$democrat)
describe(county_data$republican)
```

```{r histograms}
head(county_data %>% arrange(-pop.density),10)

hist(county_data$pop.density)
hist(county_data$college)
hist(county_data$income)
hist(county_data$democrat)
hist(county_data$republican)
```

```{r boxplots}
head(county_data %>% arrange(-pop.density),10)

boxplot(county_data$pop.density,ylab="Population Density")
boxplot(county_data$college,ylab="% with Bachelor Degree")
boxplot(county_data$income,ylab="Median Family Income ")
boxplot(county_data$democrat,ylab="% Residents Voted for Democrat")
boxplot(county_data$democrat,ylab="% Residents Voted for Republican")
```

```{r scatterplot}
ggplot(county_data,aes(x=pop.density,y=republican)) + geom_point()
ggplot(county_data,aes(x=pop.density,y=democrat))+ geom_point()

ggplot(county_data,aes(x=income,y=republican)) + geom_point()
ggplot(county_data,aes(x=income,y=democrat))+ geom_point()

ggplot(county_data,aes(x=college,y=republican)) + geom_point()
ggplot(county_data,aes(x=college,y=democrat)) + geom_point()
  
```



